Bayesian Inference from Time Series: Information Loss and the Whittle Likelihood

### 2018-03-07 doc/ms.tex
Mean vector to zero

Assume a quantity has been measured at $N$ equally-spaced time points
$\boldsymbol{t} = \{0, 1, 2, ..., N-1\}$, resulting in observations
$\boldsymbol{y} = \{y_0, y_1, ..., y_{N-1}\}$.
Throughout this paper, I will specify all models in the framework of
Gaussian processes. Ultimately, this means the probability distribution for
the data $\boldsymbol{y}$ given some parameters $\boldsymbol{\theta}$ will
be a multivariate normal distribution whose mean vector $\boldsymbol{\mu}$
and covariance matrix $\mathbf{C}$ depend on $\boldsymbol{\theta}$:
\begin{align}
\boldsymbol{y} | \boldsymbol{\theta} &\sim
    \textnormal{Normal}\left(\boldsymbol{\mu}(\boldsymbol{\theta}),
                             \mathbf{C}(\boldsymbol{\theta})\right).
\end{align}
The full expression for this probability density is
\begin{align}
p(\boldsymbol{y} | \boldsymbol{\theta})
    &= \frac{1}{\sqrt{(2\pi)^N\det \mathbf{C}}}
        \exp\left[-\frac{1}{2}
                    (\boldsymbol{y} - \boldsymbol{\mu})^T
                    \mathbf{C}^{-1}
                    (\boldsymbol{y} - \boldsymbol{\mu})\right]
\end{align}
which also provides the exact likelihood function if we wish to infer
$\boldsymbol{\theta}$ from $\boldsymbol{y}$. The determinant and inverse
are typically expensive, though in some cases fast methods are available
\citep[e.g.][]{mackay, hodlr, celerite}.

>>>>

Assume a quantity has been measured at $N$ equally-spaced time points
$\boldsymbol{t} = \{0, 1, 2, ..., N-1\}$, resulting in observations
$\boldsymbol{y} = \{y_0, y_1, ..., y_{N-1}\}$.
Throughout this paper, I will specify all models in the framework of
Gaussian processes. Ultimately, this means the probability distribution for
the data $\boldsymbol{y}$ given some parameters $\boldsymbol{\theta}$ will
be a multivariate normal distribution whose mean vector $\boldsymbol{\mu}$
and covariance matrix $\mathbf{C}$ depend on $\boldsymbol{\theta}$. Throughout
this paper the mean vector will be set to zero, so dependence on the parameters
is through the covariance matrix only:
\begin{align}
\boldsymbol{y} | \boldsymbol{\theta} &\sim
    \textnormal{Normal}\left(\boldsymbol{0},
                             \mathbf{C}(\boldsymbol{\theta})\right).
\end{align}
The full expression for this probability density is
\begin{align}
p(\boldsymbol{y} | \boldsymbol{\theta})
    &= \frac{1}{\sqrt{(2\pi)^N\det \mathbf{C}}}
        \exp\left[-\frac{1}{2}
                    \boldsymbol{y}^T
                    \mathbf{C}^{-1}
                    \boldsymbol{y}\right]
\end{align}
which also provides the exact likelihood function if we wish to infer
$\boldsymbol{\theta}$ from $\boldsymbol{y}$. The determinant and inverse
are typically expensive, though in some cases fast methods are available
\citep[e.g.][]{mackay, hodlr, celerite}.
